Reproducible Workflow on a Public Cloud for Computational Fluid Dynamics

### 2019-04-16 tex/MesnardoBarba2019.tex
Add edits with @labarba

\IEEEtitleabstractindextext{%
\begin{abstract}
In a new effort to make our research transparent and reproducible by others, we developed a workflow to run computational studies on a public cloud. It uses Docker containers to create an image of the application software stack. We also adopt several tools that facilitate creating and managing virtual machines on compute nodes and submitting jobs to these nodes. The configuration files for these tools are part of an expanded "reproducibility package" that includes workflow definitions for cloud computing, in addition to input files and instructions. This facilitates re-creating the cloud environment to re-un the simulations under the same conditions.
\end{abstract}
}

@@

Universities and national laboratories spend millions of dollars to deploy and maintain on-site high-performance computing (HPC) clusters.
At the George Washington University, we have access to a university-managed HPC cluster called Colonial One.
The cluster is now $6$ years old and approaching its end-of-life.
On average, its computational resources are idle $9\%$ of the time and unavailable to users for roughly 5 days a year (due to maintenance).
Administrators of the cluster have been recently considering integrating cloud-computing platforms in the research-computing portfolio. 
Over the past decade, cloud-computing platforms have rapidly evolved, now offering solutions for scientific applications.
From a user's point of view, a cloud platform offers great flexibility (hardware and virtual machines) with instantaneous availability of infinite (in appearance) computational resources.
No more waiting time in job-submission queues!
This promises to greatly facilitate code development, debugging, and testing. Resources allocated on the cloud are released as soon as the job is done, avoiding paying for idle time.
On a public-cloud platform with a "pay-as-you-go" type of subscription---such as Microsoft Azure, Google Cloud, Amazon AWS---a user directly sees how much it costs to run a scientific application, while such information usually remains obscure to the end-user on university-managed clusters. 
Cost models that make sense for researchers, labs, and universities are still unclear. 
Yet, this information is key when making a decision to adopt a cloud workflow for research. 
We report here what we have learned from nearly two years of using cloud computing for our simulation-based workflow, with the hope that it will shed some light on the costs and benefits, and help others considering using cloud computing for their research.

@@

Scientific publications reporting computational results often lack sufficient details to reproduce the researcher's computational environment; e.g., they may miss to mention external libraries used along with the main computational code.
We have learned the hard way how even different versions of the same external library can alter the numerical results and even the scientific findings of a computational study\cite{mesnard_barba_2017}.
This section presents an overview and mini-tutorial of the workflow we developed to guarantee the highest level of reproducibility of our computational research, with the best available technology solutions. 
In the process of creating this reproducible workflow, we also evaluated the suitability of public cloud offerings by Microsoft Azure for our research computing needs. 
The tools we adopted for computing on cloud resources are specific for this provider.

@@

\begin{figure*}[b]
    \centering
    \includegraphics[width=16cm]{figures/cloud_workflow.png}
    \caption{Reproducible workflow on the public cloud provider Microsoft Azure. Our CFD software is version-controlled with Git and GitHub. We push to DockerHub a Docker image of our CFD application with all its dependencies. Azure CLI is use to configure accounts on Microsoft Azure and to upload/download data to/from an Azure Storage account. With Batch Shipyard, we create a pool on Azure Batch and run container-based simulation using our Docker image.}
    \label{fig:cloud_workflow}
\end{figure*}

@@

To run computational jobs on Microsoft Azure, we use several tools that facilitate creating and managing virtual machines on compute nodes, and submitting jobs to those nodes.
We use a service called Azure Batch that leverages Microsoft Azure at no extra cost, relieving the user from manually creating, configuring, and managing an HPC-capable cluster of cloud nodes, including virtual machines, virtual networks, job and task scheduling infrastructure.
Azure Batch works with both embarrassingly parallel workloads and tightly coupled MPI jobs (the latter being the case of our CFD software).
To use Azure Batch, we first need to configure a workspace on Microsoft Azure.
This can be done either via the Azure Portal in a web browser or from a local terminal using the open-source tool Azure CLI.\footnote{Azure CLI (version 2.0.57): \url{https://github.com/Azure/azure-cli}}
We prefer to use the command-line solution (program \texttt{az}), as it allows us to keep track of the steps taken to configure the Azure workspace (see Listing \ref{lst:az_configure}).
First, we set the Azure subscription we want to use (let's call it \texttt{reprosubscription}).
Next, we create a resource group, \texttt{reprorg}, located in this case in the East US region, which will contain all the Azure resources.
We create an Azure Storage account, \texttt{reprostorage}, in the resource group, as well as an Azure Batch account, \texttt{reprobatch}, associated to the storage account.
Finally, we create a fileshare (in this case of size $100$ GB) in the storage.

@@

Once the configuration files are written, we invoke Batch Shipyard through the \texttt{shipyard} program on our local machine.
The folder \texttt{examples/snake2d2k35/config\_shipyard} in the repository accompanying this paper contains an example YAML files to create a pool of two \texttt{NC24r} compute nodes (featuring K80 GPUs and using InfiniBand network).
Listing \ref{lst:shipyard_run} shows the commands to run in your local terminal to create a pool of compute nodes on Azure Batch and submit jobs to it.
During the pool creation (\texttt{shipyard pool add}), the Docker image of our CFD application is pulled from the registry to the virtual machines.
With \texttt{shipyard data ingress}, we can upload input files to the compute nodes.
Once created, we submit jobs to the pool (\texttt{shipyard jobs add}); the tasks for a job will start automatically.

\begin{lstlisting}[label=lst:shipyard_run,caption={Create a pool with Batch Shipyard and submit jobs to it.}]
$ cd $CLOUDREPRO/examples/snake2d2k35
$ az storage directory create --name snake2d2k35 --share-name fileshare --account-name reprostorage
$ export SHIPYARD_CONFIGDIR=config_shipyard
$ shipyard pool add
$ shipyard data ingress
$ shipyard jobs add
\end{lstlisting}

@@

\begin{table*}[b]
    \renewcommand{\arraystretch}{1.5}
    \caption{Hardware Specifications of Nodes Used on Microsoft Azure and Colonial One. On Both Platforms, MPI Applications Take Advantage of RDMA (Remote Direct Memory Access) Network with FDR InfiniBand.}
    \label{tab:hw_specs}
    \centering
    \begin{tabular}{cccccccc}
        Platform & Node & Intel Xeon CPU & \# threads & NVIDIA GPU & RAM (GiB) & SSD Storage (GiB) \\
        \hline
        \multirow{2}{*}{Azure} & \texttt{NC24r} & Dual 12-Core E5-2690v3 (2.60GHz) & 24 & 4 x K80 & 224 & 1440 \\
        & \texttt{H16r} & Dual 8-Core E5-2667v3 (3.20GHz) & 16 & - & 112 & 2000 \\
        \hline
        \multirow{2}{*}{Colonial One} & \texttt{Ivygpu} & Dual 6-Core E5-2620v2 (2.10GHz) & 12 & 2 x K20 & 120 & 93 \\
        & \texttt{Short} & Dual 8-Core E5-2650v2 (2.60GHz) & 16 & - & 120 & 93 \\
        \hline
    \end{tabular}
\end{table*}

@@

We ran point-to-point MPI benchmarks from the Ohio State University Micro-Benchmarks suite\footnote{OSU Micro-Benchmarks (version 5.6): \url{http://mvapich.cse.ohio-state.edu/benchmarks/}} on Microsoft Azure and Colonial One, to investigate performance in terms the latency and bandwidth.
The latency test is carried out in a ping-pong fashion and measures the time elapsed to get a response; the sender sends a message with a certain data size and waits for the receiver to send back the message with the same data size.
The bandwidth test measures the maximum sustained rate that can be achieved on the network; the sender sends a fixed number of messages to a receiver that replies only after receiving all of them.
The tests ran on \texttt{NC24r} nodes on Azure and \texttt{Ivygpu} nodes on Colonial One, all of them featuring a network interface for RDMA (Remote Direct Memory Access) connectivity to communicate over an InfiniBand network.
(RDMA allows direct access to a remote's memory without involving the operating system of the host and remote.)
Fig. \ref{fig:osu_benchmarks} reports the mean latencies and bandwidths obtained over $5$ repetitions on both platforms.
For small message sizes, the average latencies on Colonial One and Azure are $1.75$ and $1.95 \mu $s, respectively.
As the message size increases, the latency becomes smaller on Azure compared to Colonial One.
The maximum sustained bandwidth rates for Colonial One and Azure are on average $4.5$ and $5.6$ GB/s, respectively.
For all message sizes, a higher bandwidth rate was achieved on Azure.
Over the last few years, Microsoft Azure has improved its HPC solution to provide networking capabilities that are comparable or even better than our 6-year-old university-managed cluster.

@@

CFD algorithms often require solving linear systems with iterative methods.
For example, the projection Navier-Stokes solver implemented in our software requires the solution of a Poisson system at every time step to project the velocity field onto the divergence-free space (to satisfy the incompressibility constraint).
We investigated the time-to-solution for a three-dimensional Poisson system (obtained with a 7-point stencil central-difference scheme) on different nodes of Microsoft Azure and Colonial One.
The solution method was a Conjugate-Gradient (CG) method with a classical algebraic multi-grid (AMG) preconditioner using an exit criterion set to an absolute tolerance of $10^{-12}$.
The iterative solver ran on CPU nodes (\texttt{H16r} instances on Azure and \texttt{Short} nodes on Colonial One) using the CG algorithm from the PETSc library\cite{balay_et_al_2018} and an AMG preconditioner from Hypre BoomerAMG.
Fig. \ref{fig:poisson_benchmarks} (top) reports the mean runtimes (averaged over $5$ repetitions) to iteratively solve the system, on a uniform grid of $50$ million cells ($1000 \times 1000 \times 50$), as we increase the number of nodes in the pool.
We also solved the Poisson system with the NVIDIA AmgX library on multiple GPU devices using \texttt{NC24r} instances on Azure and \texttt{Ivygpu} nodes on Colonial One.
The Poisson system for a base mesh of $6.25$ million cells ($500 \times 500 \times 25$) was solved on a single compute node using $12$ MPI processes and $2$ GPU devices; we then doubled the mesh size as we doubled the number of nodes (keeping the same number of MPI processes and GPUs per node).
The number of iterations to reach convergence increases with the size of the system, so we normalize the runtimes by the number of iterations.
Fig. \ref{fig:poisson_benchmarks} (bottom left) shows the normalized mean runtimes ($5$ repetitions) obtained on Azure and Colonial One.
The smaller runtimes on Azure are explained by the fact that the NC-series of Microsoft Azure features NVIDIA K80 GPU devices with a higher compute capability than the K20 GPUs on Colonial One.
The bottom-right panel of the figure reports the normalized mean runtimes obtained on Azure when we load the bandwidth with a larger problem; the base mesh now contains $25$ millions cells ($1000 \times 500 \times 50$).
We observe larger variations in the time-to-solution for the Poisson system on Microsoft Azure; runtimes are more uniform on Colonial One.

@@

To study the flow around the flying snake, we developed a CFD software called PetIBM\cite{chuang_et_al_2018}, an open-source toolbox that solves the two- and three-dimensional incompressible Navier-Stokes equations using a projection method (seen as an approximate block-LU decomposition of the fully discretized equations\cite{perot_1993}) and an immersed-boundary method (IBM).
Within this framework, the fluid equations are solved over an extended grid that does not conform to the surface of the immersed body.
To model the presence of the body, the momentum equation is augmented with a forcing term that is activated in the vicinity of the immersed boundary.
This technique allows the use of simple fixed structured Cartesian grids to solve the equations.
PetIBM implements immersed boundary methods (IBMs) that fit into the projection method; in the present study, we use the IBM scheme proposed in \cite{li_et_al_2016}.
PetIBM runs on distributed-memory architectures using the efficient data structures and parallel routines from the PETSc library.
The software also implements the possibility to solve linear systems on multiple GPU devices distributed across the nodes with the NVIDIA AmgX library and our AmgXWrapper\cite{chuang_barba_2017}.
One of the requirements for reproducible computational results is to make the code available under a public license (allowing reuse and modification by others).
In that regard, PetIBM is open source, version-controlled on GitHub,\footnote{PetIBM (version 0.4): \url{https://github.com/barbagroup/PetIBM}} and shared under the permissive (non copy-left) BSD-3 clause license.

@@

We submitted a job on Azure Batch (via Batch Shipyard,) to compute the two-dimensional flow around an anatomically accurate sectional shape of the gliding snake.
The cross-section has a chord-length $c=1$ and forms a $35$-degree angle of attack with the incoming freestream flow.
The Reynolds number, based on the freestream speed, the body chord-length, the kinematic viscosity, is set to $Re=2000$.
The immersed boundary is centered in a $30c \times 30c$ computational domain that contains just over $2.9$ million cells.
The grid is kept uniform with the highest resolution in the vicinity of the body and stretched to the external boundaries with a constant ratio (see Table \ref{tab:grid_specs} for details about the grid).
For the immersed boundary, the same grid resolution than the background fluid grid was applied.
A convective condition was used at the outlet boundary while freestream conditions were enforced on the three other boundaries.
The job was submitted to a pool of two \texttt{NC24r} nodes, using $12$ MPI processes and $2$ GPU devices per node.
It completed $200,000$ time steps (i.e., $80$ time units of flow simulation with time-step size $\Delta t = 0.0004 c / U_\infty$) in just above $7$ hours (wall-clock time).

@@

\begin{figure}[!h]
    \centering
    \includegraphics[width=8cm]{figures/wz_multi_contourf.png}
    \caption{Filled contour of the vorticity field ($-5 \leq w_z c / U_\infty \leq 5$) after $20$, $44$, $45$, and $80$ time units of flow simulation with PetIBM for the snake cross-section at a $35$-degree angle of attack and Reynolds number $2000$. Vortex merging events are responsible for the change in the wake signature and the drop in the mean value of the lift coefficient.}
    \label{fig:wz_2d}
\end{figure}

@@

Although the two-dimensional simulations can give us some insights into to flow dynamics happening in the wake behind the snake, we know that at this Reynolds number ($Re = 2000$), three-dimensional structures will develop in the wake.
We submitted jobs on Azure Batch to perform direct numerical simulation of the three-dimensional flow around a snake model: a cylinder with the same anatomically accurate cross-section.
The computational domain is extended in the $z$-direction over a length of $3.2c$ and the grid now contains about $46$ million cells (with a uniform discretization in the $z$-direction; see Table \ref{tab:grid_specs}).
The grid is kept uniform in the $z$-direction with periodic boundary conditions.
The simulation ran in a Docker container on Azure Batch using Batch Shipyard.
The job was submitted to a pool of two \texttt{NC24r} nodes using $24$ MPI processes and $4$ GPU devices per node.
The task completed $100,000$ time steps ($100$ time units with a time-step size $\Delta t = 0.01 c / U_\infty$) in about $5$ days and $16$ hours.

@@

\begin{figure*}[!h]
    \centering
    \includegraphics[width=18cm]{figures/qcrit_wx_wake3d_0020.png}
    \caption{Lateral view of the isosurfaces of the Q-criterion ($Q = 1.0$) in the wake of the snake cylinder (with a $35$-degree angle of attack) at Reynolds number $2000$. The isosurfaces are colored with the streamwise vorticity ($-5 \leq w_x c / U_\infty \leq 5$). The figure was generated using the visualization software VisIt \cite{childs_et_al_2012}.}
    \label{fig:qcrit_wx_3d}
\end{figure*}

@@

For the year $2018$, Microsoft Azure granted our research lab a sponsorship in terms of cloud credits to run CFD simulations with our in-house software.
With a ``pay-as-you-go'' subscription, users can immediately see how much it costs to run a scientific application on a public cloud; such information is often hidden to end-users on university-managed clusters.
From May to December, we spent a total of $20,614$ USD to run CFD simulations, including a couple dozens of snake simulations.
(The output of those simulations is now being processed to further analyze the complex flow dynamics generated behind the snake model.)
During that period, we have been charged for data management, networking, storage, bandwidth, and virtual machines (Table \ref{tab:azure_charges}).
More than $99\%$ of the charges incurred were for the usage of virtual machines, mainly instances from the NC-series (Table \ref{tab:nc_series}).
To run PetIBM simulations of the snake model, we used the \texttt{NC24r} virtual machines to get access to NVIDIA K80 GPUs and InfiniBand networking.
For example, the two-dimensional snake run reported in the present article cost $55.4$ USD ($2$ \texttt{NC24r} instances with a hourly price of $3.96$ USD for about $7$ hours).
The three-dimensional computation cost $1077.1$ USD ($2$ \texttt{NC24r} instances for about $136$ hours).
(Note that with 3-year-reserved instances, the two- and three-dimensional snake simulations would have only cost $24.6$ and $478.1$ USD, respectively.)

@@

Microsoft Azure offers the possibility to take advantage of surplus capacity with ``low-priority'' virtual machines that are significantly cheaper than ``dedicated'' ones (\ref{tab:nc_series}).
For example, the low-priority \texttt{NC24r} instance has an hourly cost of $0.792$ USD, $5$ times cheaper than its dedicated counterpart.
However, the reader should keep in mind that low-priority may not be available for allocation or may be preempted at any time.
In fact, low-priority instances should be avoided for long-running MPI jobs in pools where inter-node communication is enabled.
We only used dedicated virtual machines to run our CFD software.
Moreover, our job tasks used a shared filesystems (GlusterFS on compute nodes) for I/O operations and Batch Shipyard will fail to create the pool if low-priority nodes are requested.

@@

\IEEEPARstart{E}{xploring} public cloud services as an alternative (or complementary) solution to university-managed HPC clusters to run CFD simulations with our in-house software, we used Microsoft Azure to generate reproducible computational results.
Our reproducible workflow makes use of Docker containers to faithfully port the runtime environment of our CFD application to Azure,
and the command-line utilities Azure CLI and Batch Shipyard to run multi-instance tasks within Docker containers on the Azure Batch platform.
We ran CFD simulations on instances of the NC-series that feature NVIDIA K80 GPU devices and have access to InfiniBand network.
Latency and bandwidth micro-benchmarks show that our university-managed HPC system and Microsoft Azure can produce comparable runtimes.
Thanks to a Microsoft Azure sponsorship, we were able to run dozens of CFD simulations of a gliding snake's model.
Our study of the three-dimensional flow around a cylinder with an anatomically accurate cross-section of the flying snake is currently ongoing. 
We plan to analyze the output data from these runs to be published in a later paper.

>>>>

\IEEEtitleabstractindextext{%
\begin{abstract}
In a new effort to make our research transparent and reproducible by others, we developed a workflow to run computational studies on a public cloud. It uses Docker containers to create an image of the application software stack. We also adopt several tools that facilitate creating and managing virtual machines on compute nodes and submitting jobs to these nodes. The configuration files for these tools are part of an expanded "reproducibility package" that includes workflow definitions for cloud computing, in addition to input files and instructions. This facilitates re-creating the cloud environment to re-run the computations under the same conditions.
\end{abstract}
}

@@

Universities and national laboratories spend millions of dollars to deploy and maintain on-site high-performance computing (HPC) clusters.
At the George Washington University, we have access to a university-managed HPC cluster called Colonial One.
The cluster is now $6$ years old and approaching its end-of-life.
On average, its computational resources are idle $9\%$ of the time and unavailable to users for roughly 5 days a year (due to maintenance).
Administrators of the cluster have been recently considering integrating cloud-computing platforms in the research-computing portfolio. 
Over the past decade, cloud-computing platforms have rapidly evolved, now offering solutions for scientific applications.
From a user's point of view, a cloud platform offers great flexibility (hardware and virtual machines) with instantaneous availability of infinite (in appearance) computational resources.
No more waiting time in job-submission queues!
This promises to greatly facilitate code development, debugging, and testing. Resources allocated on the cloud are released as soon as the job is done, avoiding paying for idle time.
On a public-cloud platform---such as Microsoft Azure, Google Cloud, Amazon AWS---with a "pay-as-you-go" type of subscription, a user directly sees how much it costs to run a scientific application, while such information usually remains obscure to the end-user on university-managed clusters. 
Cost models that make sense for researchers, labs, and universities are still unclear. 
Yet, this information is key when making a decision to adopt a cloud workflow for research. 
We report here what we have learned from nearly two years of using cloud computing for our computational fluid dynamics (CFD) simulations, with the hope that it will shed some light on the costs and benefits, and help others considering using cloud computing for their research.

@@

Scientific publications reporting computational results often lack sufficient details to reproduce the researcher's computational environment; e.g., they may miss to mention external libraries used along with the main computational code.
We have learned the hard way how different versions of the same external library can alter the numerical results and even the scientific findings of a computational study\cite{mesnard_barba_2017}.
This section presents an overview and mini-tutorial of the workflow we developed to aim for the highest level of reproducibility of our computational research, with the best available technology solutions. 
In the process of creating this reproducible workflow, we also evaluated the suitability of public cloud offerings by Microsoft Azure for our research computing needs. 
The tools we adopted for computing on cloud resources are specific for this provider.

@@

\begin{figure*}[t]
    \centering
    \includegraphics[width=16cm]{figures/cloud_workflow.png}
    \caption{Reproducible workflow on the public cloud provider Microsoft Azure. Our CFD software is version-controlled with Git and GitHub. We push to DockerHub a Docker image of our CFD application with all its dependencies. Azure CLI is use to configure accounts on Microsoft Azure and to upload/download data to/from an Azure Storage account. With Batch Shipyard, we create a pool on Azure Batch and run container-based simulation using our Docker image.}
    \label{fig:cloud_workflow}
\end{figure*}

@@

To run computational jobs on Microsoft Azure, we use several tools that facilitate creating and managing virtual machines on compute nodes, and submitting jobs to those nodes.
We use a service called Azure Batch that leverages Microsoft Azure at no extra cost, relieving the user from manually creating, configuring, and managing an HPC-capable cluster of cloud nodes, including virtual machines, virtual networks, job and task scheduling infrastructure.
Azure Batch works with both embarrassingly parallel workloads and tightly coupled MPI jobs (the latter being the case of our CFD software).
To use Azure Batch, we first need to configure a workspace on Microsoft Azure.
This can be done either via the Azure Portal in a web browser or from a local terminal using the open-source tool Azure CLI.\footnote{Azure CLI (version 2.0.57): \url{https://github.com/Azure/azure-cli}}
We prefer to use the command-line solution (program \texttt{az}), as it allows us to keep track of the steps taken to configure the cloud workspace (see Listing \ref{lst:az_configure}).
First, we set the Azure subscription we want to use (let's call it \texttt{reprosubscription}).
Next, we create a resource group (\texttt{reprorg}) located in this case in the East US region, which will contain all the Azure resources.
We create an Azure Storage account (\texttt{reprostorage}) in the resource group, as well as an Azure Batch account (\texttt{reprobatch}) associated to the storage account.
Finally, we create a fileshare (in this case of size $100$ GB) in the storage.

@@

Once the configuration files are written, we invoke Batch Shipyard through the \texttt{shipyard} program on our local machine.
The folder \texttt{examples/snake2d2k35/config\_shipyard} in the repository accompanying this paper contains an example YAML files to create a pool of two \texttt{NC24r} compute nodes (featuring K80 GPUs and using InfiniBand network).
Listing \ref{lst:shipyard_run} shows the commands to run in your local terminal to create a pool of compute nodes on Azure Batch and submit jobs to it.
The Docker image of our CFD application is pulled from the registry to the virtual machines during the pool creation (\texttt{shipyard pool add}).
We then upload the input files to the compute nodes (\texttt{shipyard data ingress}) and submit jobs to the pool (\texttt{shipyard jobs add}). The tasks for a job will start automatically upon submission.

\begin{lstlisting}[label=lst:shipyard_run,caption={Create a pool and submit jobs to it.}]
$ cd $CLOUDREPRO/examples/snake2d2k35
$ az storage directory create --name snake2d2k35 --share-name fileshare --account-name reprostorage
$ export SHIPYARD_CONFIGDIR=config_shipyard
$ shipyard pool add
$ shipyard data ingress
$ shipyard jobs add
\end{lstlisting}

@@

\begin{table*}[b]
    \renewcommand{\arraystretch}{1.5}
    \caption{Hardware specifications of nodes used on Microsoft Azure and Colonial One. On both platforms, MPI applications take advantage of RDMA (Remote Direct Memory Access) network with FDR InfiniBand.}
    \label{tab:hw_specs}
    \centering
    \begin{tabular}{cccccccc}
        Platform & Node & Intel Xeon CPU & \# threads & NVIDIA GPU & RAM (GiB) & SSD Storage (GiB) \\
        \hline
        \multirow{2}{*}{Azure} & \texttt{NC24r} & Dual 12-Core E5-2690v3 (2.60GHz) & 24 & 4 x K80 & 224 & 1440 \\
        & \texttt{H16r} & Dual 8-Core E5-2667v3 (3.20GHz) & 16 & - & 112 & 2000 \\
        \hline
        \multirow{2}{*}{Colonial One} & \texttt{Ivygpu} & Dual 6-Core E5-2620v2 (2.10GHz) & 12 & 2 x K20 & 120 & 93 \\
        & \texttt{Short} & Dual 8-Core E5-2650v2 (2.60GHz) & 16 & - & 120 & 93 \\
        \hline
    \end{tabular}
\end{table*}

@@

We ran point-to-point MPI benchmarks from the Ohio State University Micro-Benchmarks suite\footnote{OSU Micro-Benchmarks (version 5.6): \url{http://mvapich.cse.ohio-state.edu/benchmarks/}} on Microsoft Azure and Colonial One, to investigate performance in terms the latency and bandwidth.
The latency test is carried out in a ping-pong fashion and measures the time elapsed to get a response; the sender sends a message with a certain data size and waits for the receiver to send back the message with the same data size.
The bandwidth test measures the maximum sustained rate that can be achieved on the network; the sender sends a fixed number of messages to a receiver that replies only after receiving all of them.
The tests ran on \texttt{NC24r} nodes on Azure and \texttt{Ivygpu} nodes on Colonial One, all of them featuring a network interface for RDMA (Remote Direct Memory Access) connectivity to communicate over an InfiniBand network.
(RDMA allows direct access to a remote's memory without involving the operating system of the host and remote.)
Fig. \ref{fig:osu_benchmarks} reports the mean latencies and bandwidths obtained over $5$ repetitions on both platforms.
For small message sizes, the average latencies on Colonial One and Azure are $1.75$ and $1.95 \mu $s, respectively.
As the message size increases, the latency becomes smaller on Azure compared to Colonial One.
The maximum sustained bandwidth rates for Colonial One and Azure are on average $4.5$ and $5.6$ GB/s, respectively.
For all message sizes, a higher bandwidth rate was achieved on Azure.
Over the last few years, Microsoft Azure has indeed improved its HPC solutions to provide networking capabilities that are comparable or even better than our 6-year-old university-managed cluster.

@@

CFD algorithms often require solving linear systems with iterative methods.
For example, the Navier-Stokes solver implemented in our software requires the solution of a Poisson system at every time step to project the velocity field onto the divergence-free space (to satisfy the incompressibility constraint).
We investigated the time-to-solution for a three-dimensional Poisson system (obtained with a 7-point stencil central-difference scheme) on different nodes of Microsoft Azure and Colonial One.
The solution method was a Conjugate-Gradient (CG) method with a classical algebraic multi-grid (AMG) preconditioner using an exit criterion set to an absolute tolerance of $10^{-12}$.
The iterative solver ran on CPU nodes (\texttt{H16r} instances on Azure and \texttt{Short} nodes on Colonial One) using the CG algorithm from the PETSc library\cite{balay_et_al_2018} and an AMG preconditioner from Hypre BoomerAMG.
Fig. \ref{fig:poisson_benchmarks} (top) reports the mean runtimes (averaged over $5$ repetitions) to iteratively solve the system, on a uniform grid of $50$ million cells ($1000 \times 1000 \times 50$), as we increase the number of nodes in the pool.
We also solved the Poisson system with the NVIDIA AmgX library on multiple GPU devices using \texttt{NC24r} instances on Azure and \texttt{Ivygpu} nodes on Colonial One.
The Poisson system for a base mesh of $6.25$ million cells ($500 \times 500 \times 25$) was solved on a single compute node using $12$ MPI processes and $2$ GPU devices; we then doubled the mesh size as we doubled the number of nodes (keeping the same number of MPI processes and GPUs per node).
The number of iterations to reach convergence increases with the size of the system, so we normalize the runtimes by the number of iterations.
Fig. \ref{fig:poisson_benchmarks} (bottom left) shows the normalized mean runtimes ($5$ repetitions) obtained on Azure and Colonial One.
The smaller runtimes on Azure are explained by the fact that the NC-series of Microsoft Azure features NVIDIA K80 GPU devices with a higher compute capability than the K20 GPUs on Colonial One.
The bottom-right panel of the figure reports the normalized mean runtimes obtained on Azure when we load the bandwidth with a larger problem; the base mesh now contains $25$ millions cells ($1000 \times 500 \times 50$).
We observe larger variations in the time-to-solution for the Poisson system on Microsoft Azure; runtimes are more uniform on Colonial One.

@@

To study the flow around the flying snake, we developed a CFD software called PetIBM\cite{chuang_et_al_2018}, an open-source toolbox that solves the two- and three-dimensional incompressible Navier-Stokes equations using a projection method (seen as an approximate block-LU decomposition of the fully discretized equations\cite{perot_1993}) and an immersed-boundary method (IBM).
Within this framework, the fluid equations are solved over an extended grid that does not conform to the surface of the immersed body.
To model the presence of the body, the momentum equation is augmented with a forcing term that is activated in the vicinity of the immersed boundary.
This technique allows the use of simple fixed structured Cartesian grids to solve the equations.
PetIBM implements immersed boundary methods (IBMs) that fit into the projection method; in the present study, we use the IBM scheme proposed in \cite{li_et_al_2016}.
PetIBM runs on distributed-memory architectures using the efficient data structures and parallel routines from the PETSc library.
The software also implements the possibility to solve linear systems on multiple GPU devices distributed across the nodes with the NVIDIA AmgX library and our AmgXWrapper\cite{chuang_barba_2017}.
One of the requirements for reproducible computational results is to make the code available under a public license (allowing reuse and modification by others).
In that regard, PetIBM is open source, version-controlled on GitHub,\footnote{PetIBM (version 0.4): \url{https://github.com/barbagroup/PetIBM}} and shared under the permissive (non copy-left) BSD-3 clause license.
We also provide a Docker image of PetIBM on DockerHub and its Dockerfile is available in the GitHub repository of the software.

@@

We submitted a job on Azure Batch (with Batch Shipyard,) to compute the two-dimensional flow around an anatomically accurate sectional shape of the gliding snake.
The cross-section has a chord-length $c=1$ and forms a $35$-degree angle of attack with the incoming freestream flow.
The Reynolds number, based on the freestream speed, the body chord-length, the kinematic viscosity, is set to $Re=2000$.
The immersed boundary is centered in a $30c \times 30c$ computational domain that contains just over $2.9$ million cells.
The grid is uniform with the highest resolution in the vicinity of the body and stretched to the external boundaries with a constant ratio (see Table \ref{tab:grid_specs} for details about the grid).
The discretization of the immersed boundary has the same resolution as the background fluid grid.
A convective condition was used at the outlet boundary while freestream conditions were enforced on the three other boundaries.
The job was submitted to a pool of two \texttt{NC24r} nodes, using $12$ MPI processes and $2$ GPU devices per node.
It completed $200,000$ time steps (i.e., $80$ time units of flow simulation with time-step size $\Delta t = 0.0004 c / U_\infty$) in just above $7$ hours (wall-clock time).

@@

\begin{figure}[!h]
    \centering
    \includegraphics[width=8cm]{figures/wz_multi_contourf.png}
    \caption{Filled contour of the vorticity field ($-5 \leq w_z c / U_\infty \leq 5$) after $20$, $44$, $45$, and $80$ time units of flow simulation with PetIBM for the snake cross-section at a $35$-degree angle of attack and Reynolds number $2000$. Vortex merging events trigger a change in the wake signature causing the drop in the mean value of the lift coefficient.}
    \label{fig:wz_2d}
\end{figure}

@@

Although the two-dimensional simulations can give us some insights into to flow dynamics happening in the wake behind the snake, we know that at this Reynolds number ($Re = 2000$), three-dimensional structures will develop in the wake.
We submitted jobs on Azure Batch to perform direct numerical simulation of the three-dimensional flow around a snake model: a cylinder with the same anatomically accurate cross-section.
The computational domain extends in the $z$-direction over a length of $3.2c$ and the grid now contains about $46$ million cells (with a uniform discretization in the $z$-direction and periodic boundary conditions; see Table \ref{tab:grid_specs}).
The job was submitted to a pool of two \texttt{NC24r} nodes using $24$ MPI processes and $4$ GPU devices per node.
The task completed $100,000$ time steps ($100$ time units with a time-step size $\Delta t = 0.01 c / U_\infty$) in about $5$ days and $16$ hours.

@@

\begin{figure*}[!h]
    \centering
    \includegraphics[width=18cm]{figures/qcrit_wx_wake3d_0020.png}
    \caption{Lateral view of the isosurfaces of the Q-criterion ($Q = 1$) in the wake of the snake cylinder (with a $35$-degree angle of attack) at Reynolds number $2000$. The isosurfaces are colored with the streamwise vorticity ($-5 \leq w_x c / U_\infty \leq 5$). The figure was generated using the visualization software VisIt \cite{childs_et_al_2012}.}
    \label{fig:qcrit_wx_3d}
\end{figure*}

@@

For the year $2018$, Microsoft Azure granted our research lab a sponsorship (in terms of cloud credits) to run CFD simulations with our in-house software.
With a ``pay-as-you-go'' subscription, users can immediately see how much it costs to run a scientific application on a public cloud; such information is often hidden to end-users on university-managed clusters.
From May to December, we spent a total of $20,614$ USD to run CFD simulations, including a couple dozens of snake simulations.
(The output of those simulations is now being processed to further analyze the complex flow dynamics generated behind the snake model.)
During that period, we have been charged for data management, networking, storage, bandwidth, and virtual machines (Table \ref{tab:azure_charges}).
More than $99\%$ of the charges incurred were for the usage of virtual machines, mainly instances from the NC-series (Table \ref{tab:nc_series}).
To run PetIBM simulations of the snake model, we used the \texttt{NC24r} virtual machines to get access to NVIDIA K80 GPUs and InfiniBand networking.
For example, the two-dimensional snake run reported in the present article cost $55.4$ USD ($2$ \texttt{NC24r} instances with a hourly price of $3.96$ USD for about $7$ hours).
The three-dimensional computation cost $1077.1$ USD ($2$ \texttt{NC24r} instances for about $136$ hours).
(Note that with 3-year-reserved instances, the two- and three-dimensional snake simulations would have only cost $24.6$ and $478.1$ USD, respectively.)

@@

Microsoft Azure offers the possibility of taking advantage of surplus capacity with ``low-priority'' virtual machines that are substantially cheaper than ``dedicated'' ones (\ref{tab:nc_series}).
For example, the low-priority \texttt{NC24r} instance costs $0.792$ USD per hour, $5$ times cheaper than its dedicated counterpart.
The reader should keep in mind that low-priority virtual machines may not be available for allocation or may be preempted at any time.
Thus, low-priority instances should be avoided for long-running MPI jobs in pools where inter-node communication is enabled.
We only used dedicated virtual machines to run our CFD simulations.
Moreover, our job tasks used a shared filesystem (GlusterFS on compute nodes) for I/O operations and Batch Shipyard would fail to create the pool if low-priority nodes were requested.

@@

\IEEEPARstart{E}{xploring} public cloud services as an alternative (or complementary) solution to university-managed HPC clusters to run CFD simulations with our in-house software, we used Microsoft Azure to generate reproducible computational results.
Our reproducible workflow makes use of Docker containers to faithfully port the runtime environment of our CFD application to Azure,
and the command-line utilities Azure CLI and Batch Shipyard to run multi-instance tasks within Docker containers on the Azure Batch platform.
We ran CFD simulations on instances of the NC-series that feature NVIDIA K80 GPU devices and have access to InfiniBand network.
Latency and bandwidth micro-benchmarks show that our university-managed HPC system and Microsoft Azure can deliver similar performances.
Thanks to a Microsoft Azure sponsorship, we were able to run dozens of CFD simulations of a gliding snake's model.
Our study of the three-dimensional flow around a cylinder with an anatomically accurate cross-section of the flying snake is currently ongoing. 
We plan to analyze the output data from these runs, and publish them in a later paper.
